Ms 1

### 2017-10-26 ms1/ms1_display.Rmd
updating display

n_unique_bene = scan('tables/n_unique_bene.tsv')
n_million_unique_bene = round(scan('tables/n_unique_bene.tsv') / 1e6, 2)
n_pde = sum(claims_by_abx$n)
n_million_pde = round(n_pde / 1e6, 2)

@@

There are `r n_million_unique_bene` million unique beneficiaries and `r n_million_pde` million PDEs across all study years.

@@

There were `r prettyNum(n_unique_bene, big.mark=',')` unique beneficiaries and `r prettyNum(n_pde, big.mark=',')` PDEs.

@@

## Claims by age group

```{r claims_by_age}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year') %>%
  select(name, adj_change=change_str)

@@

# do claims by age group separate
# this could be baked into the analysis step
claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
                        between(.$age, 76, 85) ~ 'age76_85',
                        between(.$age, 86, 95) ~ 'age86_95')) %>%
  group_by(year, name) %>%
  summarize_at(vars(n_bene, n_claims), sum) %>%
  ungroup() %>%
  mutate(cpkp=1000*n_claims/n_bene) %>%
  select(year, name, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
  mutate_at(vars(`2011`, `2014`), round) %>%
  mutate(group='age')

@@

@@

@@

```{r approp_changes}
# read in the raw and trend data
pde_approp_table = read_tsv('tables/pde_approp_table.tsv') %>%
  (function(df) bind_rows(
    df,
    df %>% group_by(year, tier) %>%
      summarize(n=sum(n)) %>%
      mutate(antibiotic='overall')
  )) %>%
  filter(antibiotic %in% c(top_abx, 'overall')) %>%
  group_by(year, antibiotic) %>%
  mutate(f=n/sum(n)) %>%
  ungroup()

pde_inapprop_risk_2011 = pde_approp_table %>%
  # add up tiers 3 and 4
  filter(year==2011, tier > 2) %>%
  group_by(antibiotic) %>%
  summarize(risk_2011=sum(f))

adjusted_changes_firstfill = read_tsv('tables/model_abx_firstfill.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_relevel(fct_inorder(factor(antibiotic)), 'overall', after=Inf))

# compute risk ratios
pde_inapprop_trends = read_tsv('tables/pde_inapprop_trends.tsv') %>%
  filter(term=='year', antibiotic != 'Other') %>%
  mutate(estimate=exp(3*estimate), # odds ratio over the whole time period
         std.error=sqrt(3)*std.error,
         adj.p=p.adjust(p.value, method='BH'),
         ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_high=estimate+ci) %>%
  left_join(pde_inapprop_risk_2011, by='antibiotic') %>%
  # convert OR to RR (i.e., adjusted change)
  mutate_at(vars(estimate, ci_low, ci_high), function(x) risk_ratio_f(x, .$risk_2011 - 1)) %>%
  # convert to percents
  mutate_at(vars(estimate, ci_low, ci_high), function(x) (x - 1)*100) %>%
  select(antibiotic, model, pct_change=estimate, ci_low, ci_high, adj.p) %>%
  # now estimate is adjusted change in INappropriate prescribing
  mutate(antibiotic=factor(antibiotic, levels=levels(adjusted_changes_firstfill$antibiotic), ordered=TRUE))
```

```{r adjusted_changes}
adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  select(antibiotic, pct_change, ci_low, ci_high) %>%
  arrange(pct_change) %>%
  mutate(antibiotic=fct_inorder(antibiotic),
         antibiotic=fct_relevel(antibiotic, 'overall', after=Inf))

adjusted_change_levels = adjusted_changes %>%
  filter(antibiotic != 'overall') %>%
  pull(antibiotic) %>%
  as.character

@@

adjusted_changes %>%
  mutate(antibiotic=as.character(antibiotic)) %>%
  bind_rows(pde_inapprop_trends %>%
              filter(antibiotic=='overall', model=='12/3') %>%
              mutate(antibiotic='overall_inapp')) %>%
  mutate(antibiotic=factor(antibiotic, levels=c(adjusted_change_levels, 'overall_inapp', 'overall'))) %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

@@

### First fills

Bars show the adjusted change in prescribing, considering only first fills. The blue dots show the change in inappropriate prescribing (which only includes first fills).

```{r}
ggplot(adjusted_changes_firstfill, aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=model)) +
  geom_hline(yintercept=0) + 
  geom_bar(aes(color=NULL), stat='identity', fill=NA, color='black') +
  geom_errorbar(aes(color=NULL), width=0.5) +
  geom_point(dat=pde_inapprop_trends, shape=1) +
  geom_errorbar(dat=pde_inapprop_trends, width=0.25) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change from 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 35), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_classic() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

```{r}
bind_rows(
  unadjusted_changes %>% mutate(src='unadjusted, all fills'),
  adjusted_changes %>% mutate(src='adjusted, all fills'),
  adjusted_changes_firstfill %>% mutate(src='adjusted, first fills')
) %>%
  mutate(antibiotic=factor(antibiotic, levels=c(adjusted_change_levels, 'overall'))) %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high, color=src)) +
  geom_hline(yintercept=0, color='black') +
  geom_point(shape=1) +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') + ylab('change from 2011 to 2014 (%)') +
  theme_minimal()
```

@@

### Adjusted changes in inappropriate prescribing

```{r pde_approp_trends_table}
pde_inapprop_trends %>% kable
```

## Fluoroquinolones

Consumption of ciprofloxacin is slightly decreasing, and consumption of levofloxacin is increasing. How does fluoroquinolone consumption overall change?

```{r fq_table}
fq_names = claims_by_abx %>%
  filter(str_detect(antibiotic, 'floxacin')) %>%
  group_by(antibiotic) %>%
  summarize(n=sum(n)) %>%
  arrange(desc(n)) %$%
  antibiotic

fq_table = claims_by_abx %>%
  filter(antibiotic %in% fq_names) %>%
  group_by(year) %>%
  summarize(cpkp=sum(cpkp))

fq_change_pct_unadj = fq_table %>%
  spread(year, cpkp) %$%
  { (`2014`-`2011`)/`2011` } %>%
  { round(100*., 2) }
```

```{r fq_trends}
fq_trends = read_tsv('tables/model_fq.tsv') %>%
  mult_est(3) %>%
  model_ci() %>%
  select(term, pct_change, ci_low, ci_high, p.value)

fq_change_adj = fq_trends %>%
  filter(term=='year') %>% 
  mutate_at(vars(pct_change, ci_low, ci_high), function(x) round(x, 2)) %$%
  str_interp("${pct_change}% (${ci_low}% -- ${ci_high}%)")
```

Overall, fluoroquinolones changed by a tiny amount, `r fq_change_pct_unadj`% (unadjusted). The adjusted change is `r fq_change_adj` (%) over this period. This included `r fq_names`.

@@

## Trends in appropriateness

```{r approp_graph}
pde_approp_table %>%
  #filter(year %in% c(2011, 2014)) %>%
  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
  geom_bar(stat='identity') +
  facet_grid(~antibiotic) +
  theme_classic()
```

```{r pde_approp_table}
pde_approp_table %>%
  mutate(approp=tier <= 2) %>%
  group_by(approp) %>%
  summarize(n=sum(n)) %>%
  mutate(f=n/sum(n)) %>%
  kable
```

I'll focus on tier 1 and 2 because the SSTI drugs don't have common tier 1 uses. I'll call this "high tier" prescribign. If prescribing had changed in an optimal way, only low-tier prescribing would have declined, or only high-tier prescribing would have been added.

If a drug declined by $X$, then the optimal change in the high tier fraction would be $1/(1-X) = 1+X + \mathcal{O}(X^2)$. For example, overall prescribing fell by about 5%, and high-tier prescribing increased by 2.5%. Azithromycin prescribing fell by about 20%, and high-tier azithro prescribing rose by only 6%.

I'm not sure what to make of these results, because they depend strongly on how you define the tiers. If you use only tier 1 to describe "appropriate", then levofloxacin has the biggest decline in appropriateness, but if you use tiers 1 and 2, then it's change isn't significant.

@@

For each drug, I asked what fraction of PDEs are associated with each diagnosis. This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc. The values are percents in 2011.

@@

```{r dx_to_pde}
dx_to_pde_unadj = read_tsv('tables/dx_to_pde_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  select(dx_cat, antibiotic, year, f) %>%
  spread(year, f)

>>>>

@@

@@

```{r unique_bene_pde}
round_millions = function(x) round(x / 1e6, 2)
pn = function(x) prettyNum(x, big.mark=',')

n_unique_bene = scan('tables/n_unique_bene.tsv')

claims_by_fill = read_tsv('tables/claims_by_fill.tsv')
n_pde = sum(claims_by_fill$n)
n_refills = claims_by_fill %>% filter(!first_fill) %$% n
```

There were `r round_millions(n_unique_bene)` million unique beneficiaries. They had `r round_millions(n_pde)` million PDEs, of which `r round_millions(n_refills)` million were refills.

@@

## Claims by demography

```{r claims_by_demography}
change_demog = read_tsv('tables/model_overall_demography.tsv') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  model_ci() %>%
  filter(term=='year') %>%
  select(name, adj_change=change_str)

@@

# do claims by age group separate
# this could be baked into the analysis step
claims_by_age = read_tsv('tables/claims_by_age.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(name=case_when(between(.$age, 65, 75) ~ 'age65_75',
                        between(.$age, 76, 85) ~ 'age76_85',
                        between(.$age, 86, 95) ~ 'age86_95',
                        .$age >= 96 ~ 'age96_')) %>%
  group_by(year, name) %>%
  summarize_at(vars(n_bene, n_claims), sum) %>%
  ungroup() %>%
  mutate(cpkp=1000*n_claims/n_bene) %>%
  select(year, name, cpkp) %>%
  spread(year, cpkp) %>%
  mutate(pct_change=round((`2014`-`2011`)/`2011`*100, 2)) %>%
  mutate_at(vars(`2011`, `2014`), round) %>%
  mutate(group='age')

@@

That last table could also use an "overall" row:

```{r}
claims_overall %>%
  filter(year %in% c(2011, 2014)) %>%
  mutate(cpkp=round(cpkp)) %>%
  kable
```

@@

### Trends in appropriateness

```{r tier_graph}
pde_tier_table = read_tsv('tables/pde_approp_table.tsv') %>%
  group_by(year, antibiotic) %>%
  mutate(f=n/sum(n)) %>%
  ungroup()

pde_tier_table %>%
  filter(antibiotic %in% top_abx) %>%
  ggplot(aes(x=factor(year), y=f, fill=factor(tier))) +
  geom_bar(stat='identity') +
  facet_grid(~antibiotic) +
  theme_classic()
```

```{r approp_trends}
raw_adjusted_changes = read_tsv('tables/model_abx.tsv') %>%
  rename(antibiotic=name) %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error)

# fraction inappropriate in 2011
f_i = pde_tier_table %>%
  filter(year==2011) %>%
  mutate(inapp=tier==3) %>%
  group_by(inapp) %>%
  summarize(n=sum(n)) %>%
  mutate(f=n/sum(n)) %>%
  filter(inapp) %$%
  f

# fold change in consumption
pde_trend = raw_adjusted_changes %>%
  filter(antibiotic=='overall') %>%
  # multiplications by 3 done above
  mutate(ci=1.96*std.error,
         rr=exp(estimate),
         rr_lo=exp(estimate-ci),
         rr_hi=exp(estimate+ci))

d_c = pde_trend$rr
d_c_lo = pde_trend$rr_lo
d_c_hi = pde_trend$rr_hi

pde_inapprop_trend = read_tsv('tables/pde_inapprop_trend.tsv') %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error,
         ci=1.96*std.error,
         or=exp(estimate),
         or_lo=exp(estimate-ci),
         or_hi=exp(estimate+ci))

d_f = risk_ratio_f(pde_inapprop_trend$or, f_i)
d_f_lo = risk_ratio_f(pde_inapprop_trend$or_lo, f_i)
d_f_hi = risk_ratio_f(pde_inapprop_trend$or_hi, f_i)

rr2pct = function(x) round((x - 1) * 100, 2)

inapprop_changes = data_frame(
  antibiotic=c('overall_inapp', 'overall_app'),
  pct_change=c(d_c * d_f, (1.0 - d_f * f_i) / (1.0 - f_i) * d_c),
  ci_low=c(d_c_lo * d_f_lo, (1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo),
  ci_high=c(d_c_hi * d_f_hi, (1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)
) %>%
  mutate_if(is.numeric, rr2pct)
```

Inappropriate prescribing changed by `r rr2pct(d_c * d_f)`% (95% CI `r rr2pct(d_c_lo * d_f_lo)` -- `r rr2pct(d_c_hi * d_f_hi)`). Appropriate prescribing changed by `r rr2pct((1.0 - d_f * f_i) / (1.0 - f_i) * d_c)`% (95% CI `r rr2pct((1.0 - d_f_hi * f_i) / (1.0 - f_i) * d_c_lo)` -- `r rr2pct((1.0 - d_f_lo * f_i) / (1.0 - f_i) * d_c_hi)`).

@@

```{r adjusted_changes}
adjusted_changes = raw_adjusted_changes %>%
  model_ci() %>%
  arrange(desc(pct_change)) %>%
  select(antibiotic, pct_change, ci_low, ci_high)

adjusted_change_levels = adjusted_changes %>%
  filter(antibiotic != 'overall') %$%
  { c('overall', 'overall_app', 'overall_inapp', antibiotic) } %>%
  rev

@@

adjusted_changes %>%
  bind_rows(inapprop_changes) %>%
  mutate(antibiotic=as.character(antibiotic)) %>%
  mutate(antibiotic=factor(antibiotic, levels=adjusted_change_levels)) %>%
  ggplot(aes(x=antibiotic, y=pct_change, ymin=ci_low, ymax=ci_high)) +
  geom_bar(stat='identity', fill=NA, color='black') +
  geom_errorbar(width=0.5) +
  coord_flip() +
  xlab('') +
  ylab('adjusted change in consumption 2011 to 2014 (%)') +
  scale_y_continuous(limits=c(-22, 32), breaks=c(-20, -10, 0, 10, 20, 30)) +
  theme_minimal() +
  theme(axis.title.x = element_text(size=16),
        axis.text = element_text(size=16))
```

@@

@@

@@

@@

For each drug, I asked what fraction of PDEs are associated with each diagnosis. This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc. The values are percents in 2011. A row can add up to more than 100% because one PDE can be associated with multiple diagnoses.

@@

```{r dx_to_pde}
dx_to_pde_unadj = read_tsv('tables/dx_to_pde_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  select(dx_cat, antibiotic, year, f=f_encounters_with_abx) %>%
  spread(year, f)
